### 2025-05-06 evaluating-prediction-accuracy.qmd
20250505 - predictive R-squared

###### Predictive $R^2$ {#sec-predictiveRsquared}

Predictive $R^2$ is described by @Hopper2014 and @Tuzen2025: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>) and <https://mfatihtuzen.netlify.app/posts/2025-04-30_rsquared/> (archived at <https://perma.cc/EPP8-VG23>).
Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
Larger values indicate greater accuracy.
The formula for predictive $R^2$ is in in @eq-predictiveRSquared:

$$
\begin{aligned}
  R^2_{pred} &= 1 - \frac{\text{PRESS}}{\text{SS}_\text{total}}
\end{aligned}
$$ {#eq-predictiveRSquared}

@@

For example, here is an example of estimating predictive $R^2$ 10-fold cross-validation using the `caret` package [@R-caret].

>>>>

###### Predictive $R^2$ ($R^2_{\text{pred}}$) {#sec-predictiveRsquared}

Predictive $R^2$ ($R^2_{\text{pred}}$) is described by @Hopper2014 and @Tuzen2025: <https://tomhopper.me/2014/05/16/can-we-do-better-than-r-squared/> (archived at <https://perma.cc/BK8J-HFUK>) and <https://mfatihtuzen.netlify.app/posts/2025-04-30_rsquared/> (archived at <https://perma.cc/EPP8-VG23>).
Predictive $R^2$ penalizes [overfitting](#sec-overfitting), unlike traditional $R^2$.
Larger values indicate greater accuracy.
The formula for predictive $R^2$ is in in @eq-predictiveRSquared:

$$
\begin{aligned}
  R^2_{\text{pred}} &= 1 - \frac{\text{PRESS}}{\text{SS}_\text{total}}
\end{aligned}
$$ {#eq-predictiveRSquared}

@@

For example, here is an example of estimating predictive $R^2$ using 10-fold cross-validation using the `caret` package [@R-caret].
